---
title: "package_vignette"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{package_vignette}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  warning = FALSE,
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, message = FALSE, warning=FALSE}
library(GCAMUSAJobs)
library(dplyr)
library(ggplot2)
library(knitr)
```

# Step1: read in GCAM output
This step reads in GCAM output of interest in alternative file types.

Use a local GCAM output database 
```{r Step1, warning=FALSE}
prj <- ReadGCAM(filetype = 'db',
                        input_path = "C:/GODEEP/GCAM7/output",
                        db_name = "database_basexdbGCAM-USA_Ref",
                        scen_name = "GCAM-USA_Ref", # this needs to match the scenario name specified in the DB
                        prj_name = "mydata.dat") # you name it
```

Use a local project data
```{r, warning=FALSE}
prj <- ReadGCAM(filetype = 'prj',
                input_path = "C:/GODEEP/",
                prj_name = "package.dat")
```

Use the default data (GCAM-USA 7.1 reference) in the package 
```{r}
prj <- ReadGCAM(filetype = 'prj',
                input_path = system.file("extdata", package = "GCAMUSAJobs"),
                prj_name = "package.dat")

```

# Step2: calculate detailed power generation outcomes
This step queries power generation from GCAM output, which provides generation for all capacity in operation, and calculates the implied power generation associated with installed capacity, newly added capacity, and retired capacity. The specific types of generation (“activities”) calculated are: Capacity in operation: **running**, **installed**; Capacity addition: **additions**, **add_adj**; Capacity retirements: **retirements**, **natural_retire**, **early_retire**, **early_ret_adj**, and **ret_adj**. The "_adj" activities indicate different methods of calculation (net vs. total, see Step 4). Please refer to documentation (insert a link) for details.

The power generation output is annual, broken down by state, fuel type, and technology. This step takes about 20 seconds to execute and generate output for a given scenario.


```{r Step2}
EJ_activity <- GCAM_EJ(prj)
```


**Show a subset of the outcomes**
```{r}
head(EJ_activity %>% filter(region == "TX", Year == 2050), 10)
head(EJ_activity %>% filter(region == "CA", Year == 2050), 10)
```

# Step3: calculate detailed capacity outcomes
This step derives the underlying facility capacity by different activities based on power generation outcome in **Step 2**. Capacity outcome of this step is at annual level by state, fuel type, and technology.

```{r}
GW_activity <- GCAM_GW(EJ_activity)
```

**Show a subset of the outcomes**
```{r}
head(GW_activity %>% filter(region == "TX", Year == 2050), 10)
head(GW_activity %>% filter(region == "CA", Year == 2050), 10)
```

# Step4: calculate detailed job outcomes 
This step calculates jobs based on the capacity activity outcome from **Step 3**, providing annual jobs by job type, state, fuel type, and load segment (base, intermittent, sub-peak, peak). The function accepts up to two arguments: the first is the capacity outcome from Step 3, and the second is an optional method indicator.

When the **Total** method is used (default), capacity additions and pre-mature retirements within a region can occur simultaneously for a technology, and jobs are calculated for all capacity activities. Under the **Net** method, it assumes that capacity addition and pre-mature retirement don’t happen in the same period for a technology within the same period, and jobs are calculated based on net capacity changes.


**Default: use the *Total* method**
```{r}
JOB_activity <- GCAM_JOB(GW_activity)
```

**Specify to use the *Total* method**
```{r}
JOB_activity2 <- GCAM_JOB(GW_activity, "Total")
```

**Specify to use the *Net* method**
```{r}
JOB_activity3 <- GCAM_JOB(GW_activity, "Net")
```

**Compare outcomes with *Total* and *Net* methods**
```{r}
JOB_activity3 %>% 
  dplyr::rename(Net = value) %>% 
  dplyr::left_join(JOB_activity %>% dplyr::rename(Total = value),
                   by = c("scenario", "region", "Year", "fuel", "subsector", "job", "Units")) %>% 
  dplyr::filter(region %in% c("CA", "TX"), Year == 2025, subsector == "gas_peak_CC") %>% 
  dplyr::arrange(region) %>% 
  head(15)

```


# Step5: plot employment factor

This step plots the employment factor from the JEDI (https://www.nrel.gov/analysis/jedi/models.html) model. Please refer to the documentation (**insert a link**) for details. Each point represents the employment factor of a state.

Running the function with no argument plots employment factors of all states.

```{r, warning=FALSE, fig.width=18, fig.height=6}
PLOT_EF()
```


Specifying a state of interest plots the employment factor of the chosen state.
```{r, warning=FALSE, fig.width=18, fig.height=6}
PLOT_EF("CO")
```

# Step6: plot capacity activities 

This step plots the annual average capacity by fuel and activity that is assumed to exist throughout each GCAM model timestep (by default GCAM uses 5 year steps). Red + Green is the total capacity in operation in a year, where red bars denote the capacity in operation that is newly installed in a given year, and green denotes the capacity in operation was built in the past; Blue bars denote the capacity change due to retirement.

National level
```{r, warning=FALSE, fig.width=12, fig.height=6}
PLOT_GW(GW_activity)
```


Colorado
```{r, warning=FALSE, fig.width=12, fig.height=6}
PLOT_GW(GW_activity, "CO")
```

# Step7: plot job by fuel and types
This step plots the power sector average annual direct job by fuel and job type that are assumed to exist throughout each GCAM model timestep (by default GCAM uses 5 year steps).

National level
```{r, warning=FALSE, fig.width=12, fig.height=6}
PLOT_JOB(JOB_activity)
```

Colorado
```{r, warning=FALSE, fig.width=12, fig.height=6}
PLOT_JOB(JOB_activity, "CO")
```

# Step8: plot job by job types
This step plots the power sector average annual direct jobs by job type that are assumed to exist throughout each GCAM model timestep (by default GCAM uses 5 year step).

National level
```{r, warning=FALSE, fig.width=12, fig.height=6}
PLOT_JOB_TYPE(JOB_activity)
```

Colorado
```{r, warning=FALSE, fig.width=12, fig.height=6}
PLOT_JOB_TYPE(JOB_activity, "CO")
```

# Step9: job by state of a given year
This step maps out the state-level power sector direct job in a given year.

```{r, warning=FALSE, fig.width=16, fig.height=12}
MAP_JOB(JOB_activity, 2020)
```

```{r, warning=FALSE, fig.width=16, fig.height=12}
MAP_JOB(JOB_activity, 2050)
```

```{r, warning=FALSE, fig.width=16, fig.height=12}
MAP_JOB(JOB_activity, 2100)
```
